---
output:
  pdf_document: default
  html_document: default
---



Spotify Data is being used

Reading in the dataset below and showcasing columns via str
```{r, echo=F}
knitr::opts_chunk$set(fig.width = 7, fig.height = 4,
                      message=FALSE, tidy.opts=list(width.cutoff=60)) 

```

```{r global_options, include = FALSE}
```

```{r}
library('ggplot2') 
library('tidyverse')
songs <- read.csv('spotify_songs.csv')
str(songs)
```


## Q2 Comparing Two Variances 

Continuous variable being chosen is 'danceability'.
Categorical variable being chosen is 'playlist_genre'

Below we will showcase a jitter plot with these two variables

```{r}
ggplot(songs, aes(y=playlist_genre, x=(danceability), color=playlist_genre)) +
  geom_jitter(alpha=.6) +
  theme_light() +
  theme(legend.position = 'none')+labs(x='danceability')
```

The two values from our categorical variable selected will be 'pop' and 'rap'
These can be taken as our two samples.


```{r}
sample1 <- subset(songs, playlist_genre =='pop')$danceability
sample2 <- subset(songs, playlist_genre  =='rap')$danceability
```




Now we will list the different null hypothesis which we will test.
These are the hypothesis's for comparing our population variances.

1. $H_0: \mu_1 = \mu_2 \Rightarrow \mu_1 - \mu_2 = 0$
2. $H_0: \mu_1 - \mu_2 \geq 0$
3. $H_0: \mu_1 - \mu_2 \leq 0$




First Hypothesis Check
```{r}
var.test(sample1,sample2)
```


Since our p-value is very small, we will reject the hypothesis that
danceability in both genres are the same.





Second hypothesis check

```{r}
var.test(sample1,sample2, alternative = 'less')
```

This hypothesis is also rejected as our p-value is very small, we cannot say mean
danceability in pop is higher than rap.





Third hypothesis check

```{r}
var.test(sample1,sample2, alternative = 'greater')
```

With the p-value equal to 1, we thus cannot reject this hypothesis that the mean
danceability for rap is higher than pop.

\newpage
## Q3 Comparing two population means


Listing the three hypothesis:

1. $H_0: \mu_1 = \mu_2 \Rightarrow \mu_1 - \mu_2 = 0$
2. $H_0: \mu_1 - \mu_2 \geq 0$
3. $H_0: \mu_1 - \mu_2 \leq 0$



for our first text we know the the variances do not equal each other
```{r}
t.test(sample1,sample2, var.equal = F)

```
Given our low p-value we reject the null hypothesis of the two pop means are
equal to each other.




Now for our second hypothesis, we also had to reject this when testing the 
variance, thus:

```{r}
t.test(sample1,sample2, var.equal = F, alternative = 'less')
```

Thus given our P-val, we reject the null hypothesis that pop has a higher
danceability mean than rap.




Final hypothesis check. In our variance check we were unable to reject the null
hypothesis, thus var.equal = t.

```{r}
t.test(sample1,sample2, var.equal = T, alternative = 'greater')
# var.equal = true as thus null hypothesis when testing variance was unable to
# be rejected.
```

Thus, we cannot reject this hypothesis.




