---
title: "MSCI 253 Lab 2"
author: "Kanishk Dutta"
date: "21/05/2021"
output: pdf_document
---

For our lab exercise we will be using the Spotify Data. 

First let's read in this dataset.
I'll use head to ensure that this data is read in correctly and to take
a look at some of the column names and the data in this csv.
```{r}
library('ggplot2') 
Songs <- read.csv('spotify_songs.csv')
head(Songs)

```
# Question 1

Continuous variable being chosen: 'danceability'

We will plot the histogram for this danceability variable below.
```{r}
ggplot(Songs) + geom_histogram(aes(x = danceability,y=stat(count)), 
fill = 'Light Green', colour = 'black', alpha=1, bins =20) +
    ggtitle('Histogram of track danceability') + theme_light()
```

## 1.1

Examining the histogram for the danceability variable, it is apparent that
the graph is left skewed. More tracks have a higher levels of danceability 
rather than having low levels of danceability (which makes sense, artists will
choose to make their tracks more danceable)

Examining the tails of this distribution there is a slight irregularity in the 
thickness at the right tail. (In normal distribution tails are
to decrease uniformly ). Thinking about this it does make sense as there will 
still be a substantial number of songs with a high amount of danceability
rather than a substantial number with low danceability (left tail)

Just by looking at this histogram it would be very hard to locate the quartiles
of the plot with accuracy. If we were to estimate with just the hist plot
25%: 0.5, Median: 0.65, 75%: 0.78. But these quartiles should be calculated for
better accuracy or evaluated on quantile plots.

We can take a look at these quartiles by using the function below as our
histogram does not give us an accurate understanding on it's own:
```{r}
quantile(Songs$danceability)   # finding the quartiles
```

## 1.2

Plotting the quantile plot for the danceability variable below

Below we have plotted the histogram with the quantile (having values sorted
within prob points of 0 to 1) alongside the boxplot.

Below that is also the quantile plot with just the boxplot for a less cluttered
view.

```{r}
ggplot(Songs) + 
geom_histogram(aes(y=danceability, x = -..density..), alpha = .7, colour = 'white', bins=20,fill = 'blue') +
geom_point(aes(y=sort(danceability), x=ppoints(danceability) ), colour='firebrick', alpha = .2, size =2 ) +
geom_boxplot(aes(y=danceability,x = 1.25), width=.25)  + 
  labs(x='Probability Points', y='Danceability Quantile') + theme_light()
```

```{r}
ggplot(Songs) + 
geom_point(aes(y=sort(danceability), x=ppoints(danceability) ),colour='firebrick', alpha = .2, size =2 ) +
  geom_boxplot(aes(y=danceability,x = 1.25), width=.25)  + 
  labs(x='Probability Points', y='Danceability Quantile') + theme_light()
```
Examining the quantile plot we see that it is slightly left-skewed. This is
because there is a sparsity in points at the bottom (of y-axis) and towards the
top the points are flatter (with a slight steepness at the top, thus only 
slightly left-skewed)

It is much easier to estimate where the quartiles are using the plot as we have
the prob points on our x-axis to aid us. 
  
    The Median is at about 0.66, the first quartile at about 0.55 and the third
    quartile at approx 0.76.


## 1.3

Below we utilize the fourmoments function to calculate the fourmoments for the 
danceability variable from the spotify data.


```{r}
# adding the fourmoments function which we will use to learn more about our data
library('moments')
fourmoments <- function(rv){
  c('Mean' = mean(rv), 
    'Variance' = var(rv),
    'Skewness' = skewness(rv),
    'Kurtosis' = kurtosis(rv))
}

fourmoments(Songs$danceability)
```

Examining the skewness of this variable we have a value of -0.5044, which is 
indicating that this is a left skewed distribution. 

In terms of the thickness of the tails, we examine the Kurtosis which is of a 
value of 3.01. The Kurtosis of a normal is 3, thus this being slightly higher
this distribution is leptokurtic, having slightly thicker tails than normal.


\newpage

# Question 2

Below we have chosen the variable 'liveness' which produces a 
right skewed distribution. Below is the histogram to confirm this

```{r}
ggplot(Songs) + geom_histogram(aes(x = liveness,y=stat(count)), 
fill = 'Light Green', colour = 'black', alpha=1, bins =20) +
    ggtitle('Histogram of track liveness') + theme_light()

```

Now to use qqtest to evaluate how this variable fits within different types
of distributions.

Firstly checking the normal
```{r}
library('qqtest')
qqtest(tail(Songs$liveness,300),dist = 'normal')
```
Evaluating the normal distribution it is hard to say that this of normal dist,
at both tails the points are out of the tolerable range.

Now evaluating the chi-squared plot

```{r}
library('qqtest')
qqtest((tail(Songs$liveness,300))^2,dist = 'chi-squared')
```
Chi squared seems like a good fit, all of the points from our dataset are within
the ranges of the function. (With furthest point bleeding just at edge)

Let's continue to eval the other functions, log-normal below
```{r}
library('qqtest')
qqtest(tail(Songs$liveness,300),dist = 'log-normal')
```
Log-normal is not a great fit for our data with many datapoints outside of 
the ranges.


Finally, we can evaluate the exponential dist.
```{r}
library('qqtest')
qqtest(tail(Songs$liveness,300),dist = 'exponential')
```
The exponential (along with chi-squared) also has all of our data-points in the
ranges. Both have slight deviations from the central range, we will continue 
using the exponential.


Below we will plot our ideal quantile with the sample quantile.
We use qexp below.
```{r}
library('plotly')

series <- ((Songs$danceability))
probs    <- ppoints(series)
q.exp <- qexp(probs)
q.sample <- sort((Songs$danceability))


ggplot() +
  geom_point(aes(y = q.sample, x= probs) ,colour = 'steelblue', alpha = .2) +
  geom_line(aes(y = q.exp, x = probs),colour = "black") +
    labs(x = 'Probability Points', y= 'Sample Quantiles') 

```

Comparing the ideal line with the sample quantile line, they follow a relativiley
close relation until prob 0.6 (approx) after having a high level of deviation as
the sample does not follow the ideal exponential. This is in line with what
we saw from the qqtest as the data points were scarce and not in the central range
towards the end of the exponential quantiles.






































































































